\documentclass[12pt]{article}
\usepackage[top=1cm, bottom=2cm, left=2cm, right=2cm]{geometry} 
\usepackage{color}
\usepackage[colorlinks=true, citecolor=blue, linktocpage=true]{hyperref} % n.b. glossary colour is set in style
\usepackage{enumitem}
\usepackage[style=authoryear, uniquename=false, backend=biber]{biblatex}
\usepackage{amsmath} %maths
\usepackage{graphicx}
\usepackage{import}
\usepackage{xspace}
\usepackage{perpage}
%
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\MakePerPage{footnote}

\renewcommand{\c}{\textit{c.}\xspace}
\newcommand{\eg}{\textit{e.g.}\xspace}
\newcommand{\etc}{\textit{etc.}\xspace}
\newcommand{\ie}{\textit{i.e.}\xspace}
\renewcommand{\th}{\ensuremath{^{\mbox{\tiny th}}}\xspace}
\newcommand{\Nb}{\textit{N.b.}\xspace}
\newcommand{\nb}{\textit{n.b.}\xspace}
\newcommand{\xib}{{\boldsymbol \xi}}
\newcommand{\zerob}{{\boldsymbol 0}}
%
\newcommand{\inputData}[1]{(inputs: \ref{inputs:#1})}
\newcommand{\inputDataII}[2]{(inputs: \ref{inputs:#1}, \ref{inputs:#2})}
\newcommand{\outputData}[1]{(calibProducts: \ref{calibProducts:#1})}
\newcommand{\outputDataII}[2]{(calibProducts: \ref{calibProducts:#1}, \ref{calibProducts:#2})}
\newcommand{\inputOutputData}[1]{(inputs: \ref{inputs:#1}; calibProducts: \ref{calibProducts:#1})}
\newcommand{\calypsoData}[1]{(calypsoProducts: \ref{calypso:#1})}

\newcommand{\cameraTeam}{\textbf{Camera Team deliverable}\xspace}
%
\newcommand{\TBD}{\textit{T.B.D.}\xspace}
\newcommand{\XXX}[1]{\textbf{XXX #1}\xspace}

\makeatletter
\newcommand{\myLabel}[2]{\def\@currentlabel{#2}\label{#1}}
\makeatother

\newcommand{\appRef}[1]{App. \ref{app:#1}}
\newcommand{\secRef}[1]{Sec. \ref{sec:#1}}

\bibliography{biblio,LSSTbiblio}

\begin{document}
\title{LSST's Calibration Pipelines}
\author{Robert Lupton}
%\date{1996-03-06}
\maketitle
\tableofcontents

\section{Introduction}

This document outlines the software required to produce calibration products, and then apply them.

Appendix \ref{sec:calibrationInputDatasets} and \ref{sec:calibrationTelescopeDatasets} list the data available
to the calibration products pipelines, and datasets listed therein are referenced as \eg \inputData{monoFlat}
or \calypsoData{atmosphericAbsorption}.  The outputs are listed in \secRef{calibrationOutputDatasets}, and the
datasets listed \textit{therein} are referenced as \eg \outputData{bias}.

\section{Preparing Calibration Products}

\subsection{Biases}

Bias frames will be produced by taking the median of several-to-many bias exposures.
\inputOutputData{bias}.

The individual frames will be overscan-corrected and trimmed, and then combined using standard LSST image
stacking code.  With the LSST 2s readout, we do not expect to need to remove cosmic rays explicitly, but a
robust stacking algorithm will be required.  If the readout noise in any channels is too low (relative to the
gain) to properly sample the noise distribution, a simple fix is to add sets of $n$ (\eg 3) bias exposures
before creating the stacked image.

\subsection{Amplifier Gains}
\label{sec:gains}

We expect the camera team to deliver accurate gain values for every amplifier.  Because of
correlations induced by brighter-fatter effects, these will have to be measured in a way that is independent
of intensity (modulo the effects of non-linearity in the amplifiers; \secRef{linearity}), for example
by using $^{55}$Fe or measuring the variance of binned flats. \inputData{gain}.

We will be able to monitor the relative gain within a given CCD by demanding that the flat fields be
continuous across amplifier boundaries;  this is not possible across devices.  We may be able to use
cosmic ray muons to monitor gain drifts, but it is not yet clear what accuracy will be achievable.

\subsection{Non-linearity}
\label{sec:linearity}

We expect the camera team to deliver non-linearity curves for every amplifier
\inputOutputData{linearityCurve}, and also the level above which these non-linearity curves may be unreliable.

We care about linearity for a number of purposes:
\begin{itemize}
\item Correcting potential photometric errors.
\item Making the PSF independent of brightness level.
\item Matching background levels for different readout amplifiers.
\end{itemize}

We will be able to check these values, and their stability, in at least three ways:
\begin{itemize}
\item Looking at the flux of objects as a function of the background level.
\item Looking at PSF residuals as a function of brightness and background level.
\item Looking at the offsets between amplifier segments in flat fields taken at different light levels.
\end{itemize}

\subsection{Defect Maps}

We expect the camera team to deliver a list of:
\begin{itemize}
\item Bad pixels (\ie ones that are unusable)
\item Suspect pixels (\ie ones that should raise a flag during processing
\end{itemize}
The latter list may be empty.  \inputOutputData{defectList}

This list will be augmented as new bad pixels appear or are recognised.

\subsection{Saturation Levels}

We expect the camera team to provide, for each amplifier, the lowest charge level at which charge moves from
one pixel to a neighbour.  If necessary, they will also provide the level at which the serial register
saturates (\ie if the serial saturates at a lower level than the parallels).
\inputOutputData{saturationLevel}.

These values will be modified, if necessary, based on analysis of imaging data.

\subsection{Crosstalk}
\label{sec:crosstalk}

We expect the camera team to deliver a crosstalk matrix coupling every amplifier to every other amplifier
\inputOutputData{crosstalk}; we hope and expect that most of these matrix elements will be 0.

We can can also check the cross talk coefficients using CBP data \inputData{CBP:crosstalk}, saturated
stars' bleed trails, or cosmic rays.  The latter two measurements are likely to only be useful for
checking crosstalk coefficients, while a suitably chosen set of CBP spots would allow us to make
measurements \textit{ab initio}.

\subsection{Darks}

Dark frames will be produced by taking the median of several-to-many dark exposures, each of length c. 300s;
\inputOutputData{dark}.

The individual frames will be run through standard ISR processing (including cosmic ray removal) before
being combined;  the combination may be done using standard LSST image stacking code.  If we disable the
cosmic ray removal a much larger number of exposures will be required.

\subsection{Flat Fields}

Even with high-QE devices such as those in the LSST camera, the sensitivity of the detectors is a function of
the SED of the light illuminating them; this is made worse by any spatial structure in the transmission of the
filters or of any other optical elements in the system.  Additionally, the light received at a point in the
focal plane is the sum of the focused light and of any ghost or scattered contribution.

For photometry of objects only the focused part is of interest, while for background subtraction it is the
total incident light that matters.

We plan to apply corrections to the data for the static part of the pixel size non-uniformity.  It is
\textit{very} important that the same corrections be applied to the flat fields.

\subsubsection{Broadband and Monochromatic Flats}

We will take both broadband \inputData{broadFlat} and `monochromatic' \inputData{monoFlat} flat fields.  As
discussed below, we will use the latter to estimate the correct flat field for light with any desired SED
(including that of the night sky), and the former to correct our data for the effects of dust on the filters
and other effects that change on a short time scale.

\subsubsection{Contamination of Filters and Other Optical Elements}

Dust motes appearing on filters have an affect upon the system flat fields.  This is not expected to be
a serious effect for LSST as the beam at the filter has a diameter of \c 10cm (effective diameter \c 8cm)
so only an unexpectedly large contaminant will have a measureable effect.

However, we shall take broadband flats \inputData{broadFlat} every time we put a filter into the filter
changer, and probably as a routine part of afternoon checkout.  We will synthesise a broad-band flat from the
monochromatic flats \inputData{monoFlat} and filter transmissions \inputData{filterTransmission}, and divide the
measured by the synthetic flats; by taking contaminants to be gray (\eg opaque, but with a small covering
factor) we will use the ratio to correct the monochromatic flats for any changes since they were taken.

Another approach is to use a broadband flat taken at the same time as the monochromatic flats
as the reference image.  We will do both;  any discrepancy will be a sign of potential
drift in the filter curves (see \secRef{filterTransmission}).

\subsubsection{Flats for Background Estimation}
\label{sec:backgroundFlats}

The calibration system will produce a low-resolution spectrum of the night sky \calypsoData{nightSkySpectrum},
and we will use this to synthesise a flat field image for the sky's SED from the monochromatic dome flats
\inputData{monoFlat}.

We expect that this image will have large scale structures different from those seen in sky image
that are due to illumination gradients on the flat field screen, the screen's non-uniform BDRF, and
the fact that the screen is not at infinity.  We do not expect these gradients to cause problems in
the background subtraction, and will not discuss them further except to say that if desired we could
correct them using median night-time sky frames.

Note that we are not proposing using twilight or nighttime sky flats as they do not have the desired
SED and may cause extra problems due to polarization.

\subsubsection{Flats for Photometry}
\label{sec:photometricFlats}

The flat fields described in \secRef{backgroundFlats} are not suitable for photometry of astronomical
objects for at least two reasons;  they include indirect light, and they have the wrong SED.  We can
use the Collimated Beam Projector (CBP; \inputData{CBP}) in conjunction with the monochromatic flats
\inputData{monoFlat} and filter transmission curves \inputData{filterTransmission} to solve this.

All the CBP data will be processed using the standard LSST ISR, except that no flat fields will be applied.
We will then use standard LSST aperture photometry to measure the number of counts in each CBP spot.

\begin{itemize}
\item Estimation of the Photometric Flats at a Finite Number of Points.

  If all the spots projected by the CBP were known to have the same intensity, then the spot fluxes measured
  from a scan in wavelength \inputData{CBP:mono} would give us the relative QE at a set of points in the
  camera.  In reality the spots are not all equal (due to an imperfect mask, non-uniform illumination, and
  varying plate scale) so we need to solve for their relative intensities; we can do this by moving the spots
  around the camera, \inputData{CBP:spot}, in a manner similar to the standard processing of star flats.

  We can then correct the monochromatic spots \inputData{CBP:mono} for the gains \inputData{gain}, and arrive
  at the relative QE at a set of points in the camera, as a function of wavelength, in the absence of a
  filter.

\item Estimation of the Photometric Flats for All Pixels

  These spot data \inputData{CBP:mono} only sample one point on M1, and therefore only one path
  through the optical system.  We then use the data taken at different points in M1 \inputData{CBP:M1}
  to correct these data so that they reflect the performance of the entire optical system.  We
  could sample all of M1, but in practice we expect to use many fewer pointings.

  If we could repeat this operation putting the CBP spots down on every pixel we would have our
  desired flat field;  unfortunately this is impracticable.

  Instead, we will take the monochromatic dome flats \inputData{monoFlat} and use the known QE at the position
  of the CBP spots (without applying the filter transmission curves) to correct for ghosting.  A sketch
  of the procedure is:
  \begin{itemize}
  \item Fit a surface through the CBP values (either per-CCD or for the whole camera); a spline would
    be a reasonable choice (either the product of two 1-D splines, or a thin plate spline.  I would
    start with the former as they are easier to understand).
  \item Divide the dome flat by this surface, giving an estimate of the illumination and chip-to-chip
    correction
  \item Fit a curve to this correction, and correct the dome screen.  This should be close to the
    values derived from the CBP data (and can preserve discontinuities in the QE across chips which
    the fitted curves have a hard time following).
  \item Iterate a couple of times;  each iteration should result in a smaller and smoother correction,
    which we are therefore better able to model.
  \end{itemize}

  Repeat this operation at a suitable set of wavelengths, chosen so that the variation of these corrections as
  a function of wavelength is well captured; we now know the the relative QE for all pixels in the camera, as
  a function of wavelength, in the absence of a filter.
  
  Using the filter transmission curves \inputData{filterTransmission} we can derive the relative QE for all
  pixels in the camera for each filter at 1nm resolution; this is our monochromatic photometric flatfield.

  \Nb we are in a position to provide a model of the ghosting following this analysis in two ways:
  \begin{itemize}
  \item By analysis of the CBP data, concentrating not on the spots but on the scattered/ghost light
  \item By looking at the corrections applied to the CBP spot data to arrive at the dome screen data
  \end{itemize}
  We do not need this information to calibrate LSST, but it will provide a valuable cross-check, and
  will inform the camera and telescope teams about the state of the optical elements.

\subsubsection{Outputs}
  
The corrected monochromatic flats (appropriate for the background and photometry) will be the inputs to
pipeline processing; \outputDataII{monoFlat}{monoPhotoFlat}.

We will also employ a photometric flat appropriate to a flat-spectrum source (or other well-defined SED)
\outputData{monoPhotoFlat} absorbed by a standard atmosphere.  This is not strictly-speaking an output --- it
could be constructed on the fly --- but it is more efficient to construct it here.
\outputData{standardPhotoFlat}.  \Nb the use of a standard SED/atmosphere here is intended to make the
corrections for objects' real SEDs/atmospheric absorption smaller and with smoother spatial structure.
Its use is not intrinsic to the final photometry, and if it turns out not to be helpful it may be
dispensed with.

If, as we expect, the error due to using this standard spectrum flat is slowly varying and only dependent on a
low-resolution version of the SED (\eg a resolution sufficient to capture the photometric effects of an Ia
SNe), we will prepare a low-resolution data cube to capture the essence of \outputData{monoPhotoFlat};
\outputData{monoPhotoFlatLowRes}.

\end{itemize}

\subsection{Fringe Correction}
\label{sec:fringe}

Although the thick, deep-depletion LSST CCDs reduce the amplitude of fringing they do not remove the necessity
to handle it correctly in the z and y bands.  Fringing is a QE effect, and this QE variation coupled to night
sky line emission (mostly from OH) leads to spatial structures in the background.

The classical approach to fringing is to subtract a multiple of a fringe frame, made by taking the median
stack of a very large number of science exposures scaled by their sky intensity.  Because the lines in the
night sky spectrum vary in relative intensity during the night the structure of the
fringing will, in general, vary.

In addition, the night sky shows spatial structures in rotational temperature and intensity with wavelengths
of \c 2km (the OH emission is at \c 100 km, so wavelengths of \c 1 degree, or 4 CCDs).  With a
windspeed of 60 km/s the smoothing scale (in the wind direction) is \c 1 CCD.  These spatial structures change
on time scales of \c 10 minutes.

If a CCD is of more-or-less of constant thickness with small ($\ll 1\mu$m) variations the relative variation
of the line intensities do not significantly affect the structure of the flats as the fringe patterns from
different lines are very similar (although of different amplitude and phase).  However, at least some of the
LSST chips show 20 fringes at 1$\mu$m from centre to edge of the device, which implies that the
\textit{pattern} of the fringing will change as the line ratios change.

For small fields of view, or long integrations, where the spatial structure in the night sky brightness may be
ignored it is possible to model the variation in the fringing using a PCA decomposition of sky frames.
Unfortunately LSST will show changes in fringe patterns and intensities coming from both the spectral and
spatial variation of the sky and we are planning to use a different approach.

As described in \secRef{backgroundFlats} we will synthesise flat fields that match the night sky's SED.  In a
little more detail relevant to fringing, the calibration system will produce a spectrum of the night sky with
resolution \c 200, and this is sufficient to predict the distribution of night sky line intensities
\calypsoData{nightSkySpectrum}.  We will then use this low-resolution spectrum to construct a flat field from
the monochromatic dome flats \inputData{monoFlat} that flatten the night sky spectrum, removing small-scale
structure due to the sky emission.

Note that we assume that a single spectrum within the field of view is sufficient to constrain the
night sky spectrum, whereas in reality the (atmospheric!) gravitational waves producing spatial structure
have the ability to modify the emission spectrum.   As timescales in the atmosphere are \c 10-20 minutes
we will have many realisations of the spectrum at different points in the sky, and will be able to use
these to confirm that the effective fringe image is spatially uniform;  if it transpires that this is
not the case we will be able to construct a small number of flats that capture the variation, and
use them to remove the variation in the spatial structure.

If this sky-flattening approach works we will not need any special fringe outputs;  we should
allow for the use of some variant of fringe frames as a backup \outputData{fringeFrames}.

\subsection{Filter Transmission Curves}
\label{sec:filterTransmission}

The filter transmission curves \inputData{filterTransmission} will be provided by the camera team, but by
using monochromatic spots from the CBP with \inputDataII{CBP:filter}{CBP:leak} and without filters in the beam
\inputData{CBP:mono} we can monitor the filters for any evolution, including light leaks.

Note that we are not exploring all possible light paths through the filters; while 2.5 CBP spots per CCD
mean that we have light passing through every point on the filter they are not passing through every
patch on the filter at every angle.

In theory we can measure the filter curves using the CBP, but it'd take a long time; in full generality \c 470
exposures for each 1nm step.  If we are willing to make strong enough symmetry assumptions this can be reduced
to \c 6 exposures per wavelength step; an extension of \inputData{CBP:M1} to higher spectral resolution, and
with the filter in the beam.

\subsection{Pixel Size Effects}
\label{sec:pixelSize}

It has become clear that much of the pixel-to-pixel variation seen in flat field images is
in fact due to variations in the sizes of the pixels.  These are popularily divided into two
parts:
\begin{itemize}
\item Large scale features such as the ``tree-rings'' due to variation in the properties of the
  silicon, or perturbations in the electric field (\eg ``glowing edges'').
\item Small-scale, quasi-random, variations in the pixel sizes.
\end{itemize}

If we interpret the flat fields in terms of QE variation we will make errors measuring fluxes (although
not while estimating surface brightnesses).

The pixel size variation may be thought of as a 2-D vector field $\xib$, the offset of each corner of each
pixel from a regular grid. There are various approaches being taken to measure the sizes of the pixels.

If you can reduce the distortion
to a 1-dimensional function (\eg tree rings with circular symmetry about a point; 'glowing edges' that are
a function of distance from the edge of the sensor) then high-signal-to-noise flat fields can be used to
determine the distortions \inputData{broadFlat}.

The small-scale variations are harder.  If we are willing to make the arbitrary assumption that $\xib$ is the
gradient of a scalar we can solve for it from flat field data.  Aaron Roodman's group, working on DECam data,
makes the surprising claim that they can find $\xib$ by looking for a solution close to $\xib = \zerob$.

If such approaches fail, it may be possible to measure $\xib$ by using images of 1-D sinusoids projected
onto the CCD, either by the camera team or on the mountain using the CBP.

See \secRef{correctingPixelSizes} for a discussion of how we will use $\xib$ once it is known;
\outputData{pixelSizeMap}.

\subsection{Brighter-Fatter}
\label{sec:brighterFatter}

The measurements needed to characterise Brighter-Fatter effects are not yet clear.  Current state of the art
is to use pixel-to-pixel correlations in flat fields taken at different flux levels \inputData{broadFlat}.  It
is possible that we'll need other measurements, possibly generated by projecting masks onto the camera using
the CBP;  \outputData{brighterFatterCoeffs}.

\subsection{Atmospheric Absorption}

The calibration telescope will deliver a spectrophotometrically calibrated spectrum for a bright ($\le 8$)
star in the field of almost all 8.4m visits \inputData{starSpectrum}, covering at least the bandpass of the
filter currently in use.  These stars will be chosen to have well-known spectrum
\inputData{standardStarSpectrum} based on their photometry. \XXX{This requires more thought!}

Using atmospheric models (\eg MODTRANS, extended to include a variable spectral index for the aerosols)
we will fit these two spectra to derive current atmospheric parameters in the direction of the boresight.
We expect to make use of other atmospheric information \inputData{atmosphericData} such as barometric pressure
and $O_3$ satellite data.

We will also use known photometry in the same field \inputData{photometricStandards} to help break the
degeneracy between \eg aerosol index and gray absorption.

We do not yet know how we will make use of the time-series of atmospheric parameters, nor how we will
use this series to fill in the cases where the calibration telescope fails to keep up with the 8.4m.

\Nb It is probably not necessary to know the true spectrum of the star, as we only need to reduce the
atmospheric absorption to standard conditions using the ensemble of all star spectrophotometric data.
\calypsoData{atmosphericAbsorption}

\section{Applying Calibration Products}

\subsection{Biases}

We will fit the overscan using a suitable function in the row direction (\eg a spline), taking
due account of cosmic ray hits during readout and bleeding into the overscan as saturated
stars are clocked away from the amplifier splits.

We will then subtract the bias frames \outputData{bias}.

\subsection{Linearity}

After bias subtraction we will correct for non-linearity \outputData{linearityCurve}, marking
regions where the linearity is considered unreliable as \texttt{SUSPECT}.
\Nb this assumes that the sole
source of non-linearity is in the readout amplifier.  If there is significant non-linearity
introduced further down the signal chain -- which seems unlikely -- we may need to extend
our non-linearity corrections.

\subsection{Crosstalk}

We next correct for crosstalk, including intra- and inter-chip corrections if necessary
\outputData{crosstalk}.

\subsection{Darks}

After cross-talk correction we will subtract scaled dark frames \outputData{dark}.

\subsection{Bad Pixels and Saturation}

We next interpolate over bad pixels and bleed trails \outputDataII{defectList}{saturationLevel}.

\subsection{Flat Fields}
\label{sec:applySkyFlat}

We next flat field the data, using the flat constructed using the current night-sky SED
\outputData{monoFlat}, \calypsoData{nightSkySpectrum}.  

Note that this
flat should correct for fringing, so there is no need for any fringe subtraction.   As discussed
in \secRef{fringe} it is possible that spatial and spectral variation over the field will lead
to residual fringe structures, in which case we will also performance of classical fringe frame
subtraction.

\subsection{Pixel Size Effects}
\label{sec:correctingPixelSizes}

Given the field $\xib$ \outputData{pixelSizeMap} we may redistribute flux between pixels (it's easier to think
about this as a correction in first x then y).  Linear estimation of the flux density at the pixel boundary is
probably sufficient, although higher order weights based on a nominal PSF might be better.

This flux redistribution causes a small correlation between the pixels which we need to track
with some level of accuracy.

Note that the same corrections were applied to the flat fields, so the combination of these
two corrections leaves the uniform unstructured nature of the background undisturbed.

\subsection{Brighter-Fatter}

We then redistribute flux between pixels to correct for the brighter-fatter effect
\outputData{brighterFatterCoeffs}.  Note that this does \textit{not} cause correlations between pixels, rather
it removes the correlations induced by the electric field shifts that resulted in brighter-fatter in the first
place.

\subsection{Background Subtraction}

We now proceed with normal image processing activities, leading to photometric and astrometrically
calibrated images.  At some point (\eg after background matching) we subtract the background.

\subsection{Photometric Flattening}

We can now correct the pixels for the effects of scattered light, and for the mismatch between the
colour of the sky and that of a typical astronomical object.  This is not strictly necessary --- there
are other sets of operations that result in the same final results --- but this proposal is simple
and probably wise.

Multiply by the sky flat, as computed in \secRef{applySkyFlat}, then divide by the canonical
photometric flat \outputData{standardPhotoFlat}.

\subsection{Final Photometry}
\label{sec:finalPhotometry}

For each observation of each object we know:
\begin{itemize}
\item The QE of the device as a function of wavelength at the position of the object,
  including the effects of the filter;
\item The QE of the device as a function of wavelength assumed when flat fielding;
\item The atmospheric transmission;
\item The object's colour
\end{itemize}
(Yes, there is an apparent circularity here.  But we don't need very good colours -- a few \% is fine -- so
this is not a fundamental problem;  however the details of how we arrange the computation efficiently
requires thought).

The `proper' way to measure an object's flux is given in \appRef{properFluxMeasurement}.  We will assume
here that the flat fields are slowly varying (\ie constant on the scale of an object, which means
on the scale of a PSF for objects which require exquisite photometry).

We will:
\begin{itemize}
\item Carry out appropriate pixel operations to estimate the number of counts in the object.
\end{itemize}
The rest of the operations can be carried out at the catalogue level:
\begin{itemize}
\item Look up the the canonical SED corresponding to the object's colour
\item Correct this SED for atmospheric absorption
\item Calculate the proper flat field for this SED at the centre of the object (a scalar, $C_{obj}$)
\item Lookup the standard-SED flat field at the centre of the object (another scalar, $C_{standard}$)
\item Correct the counts by $C_{obj}/C_{standard}$
\end{itemize}

In practice
we expect that  $C_{obj}$ and $C_{standard}$ will be slowly-varying functions of position, and not
require very high spectral resolution;  we will therefore use \outputData{monoPhotoFlatLowRes}.  \Nb
this will also be needed by anyone wishing to redo the photometry for \eg SNe Ia science.

%------------------------------------------------------------------------------

\appendix

\section{Calibration Data input to Calibration Products Pipelines}
\label{sec:calibrationInputDatasets}

We will have the following datasets available:
\begin{enumerate}[label=\P\arabic*,itemsep=5pt]
  \newcommand{\addMyLabel}[2]{\myLabel{#1}{\P#2}\P#2}
   \item\label{inputs:bias} Sets of bias frames.
   \item\label{inputs:gain} Gain values for all amplifiers;  note that these are required to high
     accuracy (0.1\%), as they are used while determining the photometric flats. \cameraTeam.
   \item\label{inputs:linearityCurve}  The linearity curve for every amplifier. \cameraTeam.
   \item\label{inputs:dark} Sets of 300s dark frames.
   \item\label{inputs:crosstalk} The crosstalk matrix for every pair of amplifiers in the camera.
     \cameraTeam.
   \item\label{inputs:defectList} A list of all bad pixels in each CCD. \cameraTeam.
   \item\label{inputs:saturationLevel} The level (in electrons), for each amplifier, at which charge bleeds
     to a neighbouring pixel. \cameraTeam.
   \item\label{inputs:broadFlat} Sets of flats taken through the standard LSST filters.  We will
     need flats taken at a number of flux levels to measure brighter-fatter and check linearity.
   \item\label{inputs:monoFlat} Sets of `monochromatic' (\c 1nm) flat-field screen images taken
     without the filter in the beam.
   \item\label{inputs:CBP} Sets of \textit{C}ollimated \textit{B}eam \textit{P}rojector images.
     The proposed resolutions and steps in these datasets are preliminary.
     % NOTE that the sub-CBP items have their a, b, c, ... labels set by hand, so caveat emptor
     % if you add more or change the order.  Yes, I'm lazy.
     \begin{itemize}
     \item[]\addMyLabel{inputs:CBP:mono}{\arabic{enumi}a} Sets of CBP images scanned in wavelength at 1nm resolution every
       1nm for a fixed set of spot positions on the camera, and for fixed footprint on M1.  No filter
       should be in the beam.
     \item[]\addMyLabel{inputs:CBP:spot}{\arabic{enumi}b} Sets of CBP images scanned in wavelength at 20nm resolution every
       100nm, while rotating the CBP about a pupil to move the spot pattern around the camera for a
       fixed footprint on M1.  No filter should be in the beam.
     \item[]\addMyLabel{inputs:CBP:M1}{\arabic{enumi}c} Sets of CBP images scanned in wavelength at 20nm resolution every
       100nm for a fixed set of spot positions on the camera, and for a number of footprints on M1; the
       minimum number of footprints is \c 6 for a 30cm CBP, but in reality we will explore more pointings to
       test azimuthal symmetry. No filter should be in the beam.
     \item[]\addMyLabel{inputs:CBP:filter}{\arabic{enumi}d} Sets of CBP images scanned in wavelength at 1nm resolution every
       1nm for a fixed set of spot positions on the camera, and for fixed footprint on M1.  Repeated for
       every filter. \Nb the wavelength range for each scan need only cover the range for which the filter
       transmits appreciable light.
     \item[]\addMyLabel{inputs:CBP:leak}{\arabic{enumi}e} Sets of CBP images scanned in wavelength at 20nm
       resolution every 20nm for a fixed set of spot positions on the camera, and for fixed footprint on M1.
       Repeated for every filter.
     \item[]\addMyLabel{inputs:CBP:crosstalk}{\arabic{enumi}f} Sets of CBP images taken with a suitable
       designed sparse mask to allow us to identify and measure all crosstalk images.  The simplest
       sparse mask would have only a single spot, used to illuminate each amplifier in the camera in
       turn (but less sparse solutions are probably also possible).  The wavelengths used are unimportant,
       and there are no constraints on beam footprints on M1 or filter choice.
     \end{itemize}
   \item\label{inputs:filterTransmission} Transmission curves for all the filters as a function of position.
      \cameraTeam.
   \item\label{inputs:starSpectrum}  Spectrophotometrically-calibrated spectra for stars in
     the field of view of almost all visits.
   \item\label{inputs:standardStarSpectrum} Known spectra for bright stars in the field of view of all
     visits.
   \item\label{inputs:atmosphericData}  Externally measured parameters of the atmosphere, for
     example barometric pressure and ozone.
   \item\label{inputs:photometricStandards}  Photometric standards, of a range of colours.  GAIA is a likely
     source for these data.
\end{enumerate}

%------------------------------------------------------------------------------

\section{Calibration Telescope Data Input Calibration Products Pipelines}
\label{sec:calibrationTelescopeDatasets}

\begin{enumerate}[label=$\heartsuit$\arabic*,itemsep=5pt]
   \item\label{calypso:atmosphericAbsorption} The atmospheric absorption as a function of wavelength (at
     moderate spectral resolution) for almost each visit.
   \item\label{calypso:nightSkySpectrum} Spectrum of the night sky near the Calypso boresight, with $R
     \sim 200$.\footnote{It is not entirely clear whether these will be taken on the Calypso or 8.4m
       boresight.}
\end{enumerate}

\Nb these are clearly not raw data, but will need to be extracted from spectrographic exposures.
\XXX{Add a description of those pipelines, and of the input data that they require.}

%------------------------------------------------------------------------------

\section{Outputs from LSST's Calibration Product Pipelines/Inputs to Alert/DRP Pipelines}
\label{sec:calibrationOutputDatasets}

The Calibration Products Pipeline will generate the following datasets:
\begin{enumerate}[label=\textbf{\S}\arabic*,itemsep=5pt]
   \item\label{calibProducts:bias} Trimmed, overscan subtracted, bias frame for each CCD.
   \item\label{calibProducts:linearityCurve} Linearity curves;
     identical to \inputData{linearityCurve}, unless updated during operations.
   \item\label{calibProducts:defectList} A list of all bad pixels in each CCD;
     identical to \inputData{defectList}, unless updated during operations.     
   \item\label{calibProducts:saturationLevel} The level (in electrons), for each amplifier, at which charge
     bleeds to a neighbouring pixel;
     identical to \inputData{saturationLevel}, unless updated during operations.     
   \item\label{calibProducts:crosstalk}  The crosstalk matrix for every pair of amplifiers in the camera;
     identical to \inputData{crosstalk}, unless updated during operations.
   \item\label{calibProducts:dark} Trimmed, overscan subtracted, dark frame scaled to 1 second for each CCD.
   \item\label{calibProducts:monoFlat} Sets of `monochromatic' (\c 1nm) trimmed, overscan subtracted,
     flat-field images for each filter.  These flats will \textit{include} ghost and scattered light.
   \item\label{calibProducts:monoPhotoFlat} Sets of `monochromatic' (\c 1nm) trimmed, overscan subtracted,
     flat-field images for each filter.  These flats will \textit{exclude} ghost and scattered light.
   \item\label{calibProducts:standardPhotoFlat} The linear combination of \outputData{monoPhotoFlat}
     data weighted by a flat-spectrum source (or other defined standard SED), absorbed by a standard
     atmosphere.
   \item\label{calibProducts:monoPhotoFlatLowRes} A low-resolution (in both space and wavelength)
     version of \outputData{monoPhotoFlat}.
   \item\label{calibProducts:pixelSizeMap}  A map of the pixel-size distortions.  At worse, this will be
     a $n_{\mbox{width}}\times n_{\mbox{height}}\times 2$ datacube of floats.
   \item\label{calibProducts:brighterFatterCoeffs}  Coefficients needed to model the brighter-fatter effects.
     We hope that these are a small number of floats per CCD, but this is not yet entirely clear.
   \item\label{calibProducts:fringeFrames}  \textit{Some variety of fringe frames.  We hope that we will not
     need these, but should plan for their existence as a backup.}
\end{enumerate}

%------------------------------------------------------------------------------

\section{Flux Measurement}
\label{app:properFluxMeasurement}

To measure the object's flux we should:
\begin{itemize}
\item Look up the the canonical SED corresponding to the object's colour
\item Correct this SED for atmospheric absorption
\item Calculate the proper flat field for this SED
\item Undo the flat fielding (which assumed a standard SED and atmosphere)
\item Apply the proper flat field for this SED
\item Carry out appropriate pixel operations to estimate the number of counts in the object.
\end{itemize}

We describe a simplification of these operations in \secRef{finalPhotometry}.

%------------------------------------------------------------------------------

% \printbibliography[heading=bibintoc]

\end{document}
